Simulation

### 2023-10-31 simulation.qmd
Wrangle data for brms

```{r}
#' Simulate ar(1) data
#' 
#' Adapted from <https://experienced-sampler.netlify.app/post/stan-hierarchical-ar/#testing>
#'
#' @param I number of individuals
#' @param t number of time points per individual
#' @param sigma residual standard deviation of observations
#' @param alpha_hat intercept mean
#' @param alpha_scale intercept standard deviation
#' @param beta_hat ar(1) slope mean
#' @param beta_scale ar(1) slope standard deviation
#'
#' @return a list of variables to send to Stan
#'
#' @examples simulate()
simulate <- function(
    I = 30, 
    t = 10, 
    sigma = 1, 
    alpha_hat = 4, 
    alpha_scale = 1, 
    beta_hat = 0.4, 
    beta_scale = 0.1
) {

@@

@@

```{r}
fit <- cmdstanr::cmdstan_model("model-ar1.stan", compile = FALSE)
fit$format(
  canonicalize = list("deprecations"), 
  overwrite_file = TRUE, 
  backup = FALSE
)
```

>>>>

```{r}
#' Simulate ar(1) data
#' 
#' Adapted from <https://experienced-sampler.netlify.app/post/stan-hierarchical-ar/#testing>
#'
#' @param I number of individuals
#' @param t number of time points per individual
#' @param sigma residual standard deviation of observations
#' @param alpha_hat intercept mean
#' @param alpha_scale intercept standard deviation
#' @param beta_hat ar(1) slope mean
#' @param beta_scale ar(1) slope standard deviation
#'
#' @return a list of variables to send to Stan
#'
#' @examples simulate()
simulate <- function(
    I = 50, 
    t = 25, 
    sigma = 1, 
    alpha_hat = 4, 
    alpha_scale = 1, 
    beta_hat = 0.4, 
    beta_scale = 0.1
) {

@@

## Stan model

@@

```{r}
# model <- cmdstan_model("model-ar1.stan", compile = FALSE)
# model$format(
#   canonicalize = list("deprecations"), 
#   overwrite_file = TRUE, 
#   backup = FALSE
# )
# model$compile()
model <- cmdstan_model(exe_file = "model-ar1")
path <- "models/cmdstanr-example.rds"
if (!file.exists(path)) {
post <- model$sample(
  data = dat, 
  adapt_delta = 0.99
) |> 
  as_draws_df()
saveRDS(post, path)
} else {post <- readRDS(path)}
```

Visualize estimated and defined parameters

```{r}
post <- post |> 
  select(starts_with("beta_"), ends_with("_raw")) |> 
  rename_with(~str_remove(., "_raw"))
post |> 
  pivot_longer(everything()) |> 
  ggplot(aes(value, name)) +
  scale_x_continuous(
    "Parameter value"
  ) +
  stat_histinterval(
    normalize = "xy",
    scale = 0.25, 
    size = 1
  ) +
  geom_point(
    data = as_tibble(dat) |> 
      select(sigma:beta_scale) |> 
      distinct() |> 
      pivot_longer(everything()),
    position = position_nudge(y = -0.1),
    color = "red"
  ) +
  theme(
    axis.title.y = element_blank()
  )
```

## brms

```{r}
dat <- as_tibble(dat)
dat <- dat |> 
  mutate(yt1 = lag(y), .by = individual)
dat
```
